Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-02-04 2_SSL_EES/article/article_main.tex
various minor changes

During the years following the market introduction of white LEDs, the primary metric of progress in solid-state lighting was typically luminous flux. Progress in this metric was described by “Haitz’s Law”, an observation on the increasing brightness and reduction in retail price of devices \cite{haitz1999case}\cite{haitz2011solid}. The luminous flux (total brightness) of early white LEDs was in facto too small to allow for the economical combination of multiple LEDs into lamps for general illumination purposes. In 2000, highest performing devices yielded around 10lm, just below the output of a candle as defined in the unit candela (1cd=12.57lm) \cite{haitz2011solid}. Today, however, highest performing devices yield in excess of 1600lm, the equivalent of a 100W incandescent bulb \cite{cree2020bright}. LED brightness has become sufficient to enable the construction of lamps from multiple LED devices, with design improvements focusing on higher efficiency and quality of light instead of brightness. Even though some scientific publications and industry periodicals continue to focus on brightness as a metric of progress in lighting, we find that, by now, it fails to capture the complexity of the multitude of efficiency improvements that have been driving overall LED efficiency \cite{weinold2021compound}.

@@

The historical development of a selection of the sub-efficiencies introduced in Section \ref{sec:device_performance_metrics} is shown in Figure \ref{fgr:subeff}. Using this data, together with data on the remaining sub-efficiencies not shown in this figure, we calculated the overall LED efficiency in for four years: 2003, 2010, 2016 and 2020. Figure \ref{fgr:waterfall} shows how improvements in sub-efficiencies led to improvements in overall white LED lamp efficiency from $\eta_L=5.8\%$ in 2003 to $12.6\%$ in 2010, $32.5\%$ in 2016 and finally to $38.8\%$ in 2020. As is evident from the figure, no single loss channel dominates in terms of its contribution to the overall efficiency, in line with previous observations\cite{tsao2010solid}. We note, however, that the loss channels with a fixed physical limit, e.g., Stokes loss that contributes to the conversion efficiency by phosphors, has become more dominant in 2016 and 2020 compared to 2003 and 2010. This is a direct result of the large efficiency improvements of upstream sub-efficiencies.

Figure \ref{fgr:breakthroughs_efficiency} shows the overall magnitude of contributions of identified LED innovations and technology spillovers to improvements in LED efficiency over time across different sub-efficiencies. The list of technology spillovers identified in our study is provided in Table \ref{tab:spillovers} and discussed in Section \ref{subsec:spillovers}. Through the index decomposition analysis described in Section \ref{subsec:metrics}, we find that out of the overall LED efficiency increase of 32.9\% from 5.8\% to 38.8\% between 2003 and 2020, at least 2.8\% can be attributed specifically to technology spillovers identified in this study, corresponding to 8.5\% of the total LED efficiency improvements between 2003 and 2020.

@@

Our findings are further supported by sensitivity analysis, presented in the Supplementary Information section SI4, where we find that sensitivity of the model to variation in its main parameters decreases over time with the increase of the number of DPW. In SI section 5 we also provide a comparison of our model with past cost calculations and projections published by US DOE on the basis of the \textit{LEDCOM} cost model \cite{ledcomv2} and industry data reported to the DOE as part of SSL round tables \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid}.

@@

Our findings show that the overall white LED efficiency has increased from $\eta_L=5.8\%$ in 2003 to $\eta_L=38.7\%$ in 2020. This efficiency increase was predominantly the result of a series of innovations in the LED architecture and materials as well as changes in the manufacturing process. Our comparison of improvements in device sub-efficiencies between 2003 and 2020 shows that the overall efficiency improvement was not dominated by any single sub-efficiency channel. Instead, there has been consistent progress across all device sub-efficiencies in the recorded period. Some sub-efficiencies, such as forward voltage efficiency, internal quantum efficiency and light extraction efficiency, are now within $\sim~10\%$ of their respective physical limits. A marked trend is the increasing relative contribution of losses in channels with physical limits on efficiency due to the underlying physics, such as light conversion efficiency of phosphors determined by the Stokes shift. Spectral efficiency also remains at $\sim17\%$ from its physical limit, showing potential for further improvements. Research on improving LED performance across efficiency loss channels continues \cite{cho2017white}\cite{Weisbuch2020}.

We also find that there has been a significant progress in consumer experience metrics, which was driven by innovations in the means of conversion of blue light generated by conventional GaN LEDs into white light required for general illumination. The first commercial white LEDs produced by Nichia in 1996 used a YAG phosphor that could generate only cool white light \cite{BANDO1998}. After a series of breakthrough innovations in phosphors represented in Figure 8 and described in SI section 3, LEDs today can be tuned for high colour rendering performance, high colour saturation and a range of desirable colour temperatures.

@@

Our LED manufacturing cost model shows that the cost of manufacturing low-to-mid power GaN-based white LED packages with classical chip architecture at a U.S. location using state-of-the-art equipment has decreased by 95.5\% from 1.11\$(2020) in 2003 to 0.05\$(2020) in 2020. In contrast with LED performance improvements, where progress was driven mostly by LED technology innovations, the dramatic decline in LED manufacturing cost happened mostly due to economies of scale stemming from an increase in the wafer size and higher yield per manufacturing step. Both were enabled by advances in manufacturing equipment performance and incremental process improvements from learning-by-doing.

@@

Our study has analysed in a systematic and granular way the sources of dramatic cost reductions and performance improvements in white light-emitting diodes since their introduction to the market in 1996. We find that the total LED device efficiency increase from 5.8\% to 38.7\% between 2003 and 2020, as well as improvements in consumer experience metrics, have been predominantly driven by LED technology innovations that affected all physical energy loss channels and corresponding device sub-efficiencies. We also find that among those innovations, at least nine were driven by knowledge spillovers originating in areas of science and technology beyond LEDs or solid-state lighting. These spillover-driven innovations were responsible for 8.5\% of the total efficiency improvements and nearly 100\% of the improvements in consumer experience metrics.

>>>>

During the years following the market introduction of white LEDs, the primary metric of progress in solid-state lighting was typically luminous flux. This was because the luminous flux (total brightness) of early white LEDs was too small to allow for the economical combination of multiple LEDs into lamps for general illumination purposes. In 2000, highest performing devices yielded around 10lm, just below the output of a candle as defined in the unit candela (1cd=12.57lm)\cite{haitz2011solid}. Progress in this metric was commonly rendered together with the associated reduction in retail prices as “Haitz’s Law”\cite{haitz1999case}\cite{haitz2011solid}. Today, however, highest performing devices yield in excess of 1600lm, the equivalent of a 100W incandescent bulb \cite{cree2020bright}. LED brightness has thereby become sufficient to enable the construction of lamps from multiple LED devices, with contemporary improvements focusing instead higher efficiency and quality of light instead of brightness. Even though a large number of scientific publications and industry periodicals continue to focus on brightness as a metric of progress in lighting, we find that, by now, it fails to capture the complexity of the multitude of efficiency improvements that have been driving overall LED efficiency \cite{weinold2021compound}. Specifically,  the stagnating levels of luminous flux in highest performing devices fail to capture the majority of research, which serves to improve electrical efficiency at constant brightness.

@@

The historical development of a selection of the sub-efficiencies introduced in Section \ref{sec:device_performance_metrics} is shown in Figure \ref{fgr:subeff}. Using this data, together with data on the remaining sub-efficiencies not shown in this figure, we calculated the overall LED efficiency in for four years: 2003, 2010, 2016 and 2020. Figure \ref{fgr:waterfall} shows how improvements in sub-efficiencies led to improvements in overall white LED lamp efficiency from $\eta_L=5.8\%$ in 2003 to $12.7\%$ in 2010, $32.5\%$ in 2016 and finally to $38.8\%$ in 2020. As is evident from the figure, no single loss channel dominates in terms of its contribution to the overall efficiency, in line with previous observations\cite{tsao2010solid}. We note, however, that the loss channels with a fixed physical limit, e.g., Stokes loss that contributes to the conversion efficiency by phosphors, has become more dominant in 2016 and 2020 compared to 2003 and 2010. This is a direct result of the large efficiency improvements of upstream sub-efficiencies.

Figure \ref{fgr:breakthroughs_efficiency} shows the overall magnitude of contributions of identified LED innovations and technology spillovers to improvements in LED efficiency over time across different sub-efficiencies. The list of technology spillovers identified in our study is provided in Table \ref{tab:spillovers} and discussed in Section \ref{subsec:spillovers}. The complete list of all LED innovations is provided in Table SI 2 in Section 3 of the Supplementary Information. Through the index decomposition analysis described in Section \ref{subsec:metrics}, we find that out of the overall LED efficiency increase of 32.9\% from 5.8\% to 38.8\% between 2003 and 2020, at least 2.8\% can be attributed specifically to technology spillovers identified in this study, corresponding to 8.5\% of the total LED efficiency improvements between 2003 and 2020.

@@

Our findings are further supported by apreliminary sensitivity analysis, presented in Section 2 of the Supplementary Information section, where we find that the sensitivity of the model to variation in its main parameters decreases over time with the increase of the number of DPW. We also provide a comparison of our model with past cost calculations and projections published by the US DOE on the basis of the \textit{LEDCOM} cost model \cite{ledcomv2} and industry data reported to the DOE as part of SSL round tables \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid} in Section 2 of the Supplementary Information.

@@

Our findings show that the overall white LED efficiency has increased from $\eta_L=5.8\%$ in 2003 to $\eta_L=38.8\%$ in 2020. This efficiency increase was predominantly the result of a series of innovations in the LED architecture and materials as well as changes in the manufacturing process. Our comparison of improvements in device sub-efficiencies between 2003 and 2020 shows that the overall efficiency improvement was not dominated by any single sub-efficiency channel. Instead, there has been consistent progress across all device sub-efficiencies in the recorded period. Some sub-efficiencies, such as forward voltage efficiency, internal quantum efficiency and light extraction efficiency, are now within $\sim10\%$ of their respective physical limits. A marked trend is the increasing relative contribution of losses in channels with physical limits on efficiency due to the underlying physics, such as light conversion efficiency of phosphors determined by the Stokes shift. Spectral efficiency also remains at $\sim17\%$ from its physical limit, showing potential for further improvements. Research on improving LED performance across efficiency loss channels continues\cite{cho2017white}\cite{Weisbuch2020}.

We further find that there has been significant progress in performance across the set of consumer experience metrics, which was driven by innovations related to the conversion of blue light generated by conventional GaN LEDs into the white light required for general illumination. The first commercial white LEDs produced by Nichia in 1996 used a YAG phosphor that could generate only cool white light \cite{BANDO1998}. After a series of breakthrough innovations in phosphors represented in Figure \ref{fgr:consumer_experience} and described in Section 4 of the Supplementary Information, LEDs today can be tuned for high colour rendering performance, high colour saturation and a range of desirable colour temperatures.

@@

Our LED manufacturing cost model shows that the cost of manufacturing low-to-mid power GaN-based white LED packages with classical chip architecture at a U.S. location using state-of-the-art equipment has decreased by 95.5\% from 1.11\$(2020) in 2003 to 0.05\$(2020) in 2020. In contrast with LED performance improvements, where progress was driven mostly by LED technology innovations, the dramatic decline in LED manufacturing cost resulted from an increase in the wafer size used in manufacturing and higher yields across manufacturing steps. Both were enabled by advances in manufacturing equipment performance and incremental process improvements from learning-by-doing.

@@

Our study has analysed the sources of the dramatic cost reductions and performance improvements in white light-emitting diodes since their introduction to the market in 1996 in a systematic and granular way. We find that the total LED device efficiency increase from 5.8\% to 38.8\% between 2003 and 2020, as well as improvements in consumer experience metrics, have been predominantly driven by LED technology innovations that affected all physical energy loss channels and corresponding device sub-efficiencies. We also find that among those innovations, at least nine were driven by knowledge spillovers originating in areas of science and technology beyond LEDs or solid-state lighting. These spillover-driven innovations were responsible for 8.5\% of the total efficiency improvements and nearly 100\% of the improvements in consumer experience metrics.
